---
title: "Case study and analysis of Global GDP data - 2012"
author: "Team: Vivek Viswanathan, Vanessa Torres, Gregory Asamoah, Kevin Mendonsa"
date: "Date: October 23, 2017"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Analysis of Global GDP data for 2012

The introduction should not start with "For my project I .". The introduction needs to be written as if you are presenting the work to someone who has given you the data to analyze and wants to understand the result. In other words, pretend it's not a case study for a course. Pretend it's a presentation for a client.

## R - environment

```{r environment, echo=FALSE}
sessionInfo()
```

## R- Packages used in the analysis of the data for this study

You will need to install the following packages and load them.

```{r install_load_packages}
# Install packages required for data analysis
# install.packages("tidyverse")
# install.packages("repmis")
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("ggplot2")

# Load libraries required
library(repmis)
library(dplyr)
library(tidyverse)
library(tidyr)
library(ggplot2)
```


## Please set the "BaseDir path" to reflect your own working directory


```{r set_working_directory}
BaseDir <- "C:/Vivek/Data_Science/MSDS6306-DoingDataScience/CaseStudy1/MSDS6306-Unit6-CaseStudy"
DataDir <- paste(BaseDir,"Data", sep = "/")
SourceDir <- paste(BaseDir,"Source", sep = "/")
PaperDir <- paste(BaseDir,"Paper", sep = "/")

# set working directory
# setwd(DataDir)
```


## Let's begin by downloading the data files for study

```{r data_acquisition}
# URL for file download
DownloadUrl <- "https://d396qusza40orc.cloudfront.net"

# Set the file path in the website
GDPRemoteFile<- paste(DownloadUrl, "getdata%2Fdata%2FGDP.csv", sep = "/")
CountryRemoteFile <- paste(DownloadUrl, "getdata%2Fdata%2FEDSTATS_Country.csv", sep = "/")

# Set the file paths in the local machine
GDPLocalFile<- paste(DataDir, "getdata%2Fdata%2FGDP.csv", sep = "/")
CountryLocalFile <- paste(DataDir, "getdata%2Fdata%2FEDSTATS_Country.csv", sep = "/")

# Download the files from Remote to Local
download.file(GDPRemoteFile, GDPLocalFile)
download.file(CountryRemoteFile, CountryLocalFile)

# load data files to variables and replace all "empty" strings with "NA"
GDPDataFull <- read.csv(GDPLocalFile, header=T, quote = "\"", skip=4, na.strings = c("", "NA"))
CountryData <- read.csv(CountryLocalFile, header=T, na.strings = c("", "NA"))

```


## Tidying the data to enable answering questions of key interest

```{r data_tidying}
# Set Column Names for GDPData 
names(GDPDataFull) <- c("CountryCode", "Ranking", "X.1", "Long.Name", "GDPinMillionsofDollars", "X.2","X.3", "X.4", "X.5", "X.6")
RegionData <- slice(GDPDataFull,(219:231))
WorldData <- slice(GDPDataFull, 217)
GDPDatawoRegion <- slice(GDPDataFull, 1:215)

# Cleanse the WorldData
WorldData <- select(WorldData, CountryCode, Ranking, Long.Name, GDPinMillionsofDollars) %>%
  mutate(GDPinMillionsofDollars=gsub(",","",GDPinMillionsofDollars)) %>%
  mutate(GDPinMillionsofDollars=as.numeric(GDPinMillionsofDollars))

# Cleanse the RegionData
RegionData <- select(RegionData, CountryCode, Ranking, Long.Name, GDPinMillionsofDollars) %>%
  mutate(GDPinMillionsofDollars=gsub(",","",GDPinMillionsofDollars)) %>%
  mutate(GDPinMillionsofDollars=as.numeric(GDPinMillionsofDollars))

# Cleanse the GDPDataFull
GDPDataFull <- select(GDPDataFull, CountryCode, Ranking, Long.Name, GDPinMillionsofDollars) %>%
  mutate(GDPinMillionsofDollars=gsub(",","",GDPinMillionsofDollars)) %>%
  mutate(GDPinMillionsofDollars=as.numeric(GDPinMillionsofDollars)) %>%
  mutate(Ranking=as.numeric(Ranking))

# Cleanse the GDPDatawoRegion
GDPDatawoRegion <- select(GDPDatawoRegion, CountryCode, Ranking, Long.Name, GDPinMillionsofDollars) %>%
  mutate(GDPinMillionsofDollars=gsub(",","",GDPinMillionsofDollars)) %>%
  mutate(GDPinMillionsofDollars=as.numeric(GDPinMillionsofDollars)) %>%
  mutate(Ranking=as.numeric(Ranking))
```


## KEY QUESTIONS OF INTEREST FOR GLOBAL GDP DATA

## Q1. How many of the country IDs match?

```{r key_question1}
# Merge the Data Sets based on the country shortcode or id.

CombinedGDPFullCountry <- merge(GDPDataFull, CountryData, by = c("CountryCode"))

# 224 IDs match on the Merged Data when GDP File included the Region Information.

CombinedGDPwoRegionCountry <- merge(GDPDatawoRegion, CountryData, by = c("CountryCode"))

# 210 IDs match on the Merged Data when GDP File included the Region Information.
```
## 224 records had a match on the Merged Data when the GDP File included the Region data.

## 210 records had a match on the Merged Data when the GDP File excluded the Region data.
##     This may not be accurate..I think it 190...need to revalidate

## -----------------------------------------------------------------------------------------------------------


## Q2. What is the 13th country in the resulting data frame?

```{r key_question2}
# Sort the data frame in ascending order by GDP (so United States is last).
# Sort Combined Data by GDPinMillionsofDollars

CombinedGDPFullCountry <- arrange(CombinedGDPFullCountry, CombinedGDPFullCountry$GDPinMillionsofDollars)
CombinedGDPwoRegionCountry <- arrange(CombinedGDPwoRegionCountry, CombinedGDPwoRegionCountry$GDPinMillionsofDollars)
```

## St. Kitts and Nevis is the 13th country in both the data frames, i.e., the dataset including the region and the dataset excluding the region data.

## -----------------------------------------------------------------------------------------------------------


## Q3. What are the average GDP rankings for the "High Income: OECD" and "High Income: nonOECD" groups?

```{r key_question3}
select(CombinedGDPFullCountry, Income.Group, GDPinMillionsofDollars) %>%
  group_by(Income.Group) %>%
  summarise(avg = mean(GDPinMillionsofDollars, na.rm=T))
```

## The average GDP of High Income: nonOECD is USD$   104,349.83
## The average GDP of High Income:    OECD is USD$ 1,483,917.13

## -----------------------------------------------------------------------------------------------------------


# Q4:	Plotting the GDP for all of the countries. 

```{r key_question4, echo=FALSE}
# Using ggplot2 to color the plot by Income Group.
ggplot(data=CombinedGDPwoRegionCountry, mapping = aes(x=CountryCode, y=GDPinMillionsofDollars)) +
  geom_point(mapping = aes(color =Income.Group)) +
  theme(axis.text.x=element_text(angle=90,hjust=1)) 
```

## -----------------------------------------------------------------------------------------------------------


## Q5: How many countries are Lower middle income but among the 38 nations with highest GDP?

```{r key_question5}
# Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group.
summary(CombinedGDPwoRegionCountry$Ranking)

count(filter(CombinedGDPwoRegionCountry, Ranking > 143 & Income.Group =='Lower middle income'))
```

## There are 17 countries among the 38 nations with the Highest GDP, but in the Lower middle income category.



